### 2022-09-06 doc/src/AdvPLIC.tex
Numerous small improvements to the text

In a {\RISCV} system, a Platform-Level Interrupt Controller (PLIC)
handles external interrupts that are signaled through wires rather than
by MSIs.
When the {\RISCV} harts in a system do not have IMSICs, the harts
themselves do not support MSIs, and all external interrupts to such
harts must pass through a PLIC.
But even in machines where harts have IMSICs and most interrupts are
communicated via MSIs, it is not unusual for some device interrupts
still to be signaled by dedicated wires.
In particular, for devices (or device controllers) that do not
otherwise need to initiate bus transactions
in the system, the cost of supporting
MSIs is especially high, so wired interrupts are a frugal alternative.
Wired interrupts also continue to be universally supported by all
current computer platforms, unlike MSIs, making another reason for many
commodity devices or controllers to choose wired interrupts over MSIs,
when not implementing a standard like PCI Express that dictates MSIs.

@@

\begin{commentary}
Although an APLIC with multiple interrupt domains may appear to duplicate
the per-source state listed above (source configuration, etc.) by a
factor equal to the number of domains, in fact, APLIC implementations
can exploit the fact that each source is ultimately active in only one
domain.
In all domains to which a specific interrupt source has not been
delegated, the state associated with the source appears as read-only
zeros, requiring no physical register bits.
\end{commentary}

@@

If the interrupt domain's harts have IMSICs, the relevant interrupt
files of those IMSICs must support value \z{0x40000000} for register
\z{eidelivery}, else setting DM to zero (direct delivery mode) will
always have the same effect as setting IE to zero.
See Sections \ref{sec:IMSIC-reg-eidelivery}
and~\ref{sec:AdvPLIC-directMode-intrDelivery}.

@@

An interrupt source is inactive in the interrupt domain if either the
source is delegated to a child domain (D~=~1) or it is not delegated
(D~=~0) and SM is Inactive.
Whenever interrupt source~$i$ is inactive in an interrupt domain,
the corresponding interrupt-pending and interrupt-enable bits within
the domain are read-only zeros, and register \z{target[$i$]} is also
read-only zero.
If source~$i$ is changed from inactive to an active mode, the
interrupt source's pending and enable bits remain zeros, unless set
automatically for a reason specified later in this section or in
Section~\ref{sec:AdvPLIC-pendingBits}, and \z{target[$i$]} becomes an
{\unspecified} valid value.

@@

It is expected that most systems will arrange the
physical addresses of target IMSICs in a simple
linear correspondence with hart index numbers.
Registers \z{mmsiaddrcfg} and \z{mmsiaddrcfgh} (along with
\z{smsiaddrcfg} and \z{smsiaddrcfgh} from the next subsection) allow
sufficiently trusted machine-level software, early after system reset,
to configure the pattern of physical addresses for target IMSICs and
then lock this configuration against subsequent tampering.

@@

@@

%--------------------
\subsubsection*{Effect of changing delivery mode}

If field DM of register \z{domaincfg} is changed, the \z{target}
registers for all active interrupt sources within the domain become
{\unspecified} valid values.

>>>>

In a {\RISCV} system, a Platform-Level Interrupt Controller (PLIC)
handles external interrupts that are signaled through wires rather than
by MSIs.
When the {\RISCV} harts in a system do not have IMSICs, the harts
themselves do not support MSIs, and all external interrupts to such
harts must pass through a PLIC.
But even in machines where harts have IMSICs and most interrupts are
communicated via MSIs, it is not unusual for some device interrupts
still to be signaled by dedicated wires.
In particular, for devices (or device controllers) that do not
otherwise need to initiate bus transactions
in the system, the cost of supporting
MSIs is especially high, so wired interrupts are a frugal alternative.
Wired interrupts also continue to be universally supported by all
current computer platforms, unlike MSIs, making another reason for many
commodity devices or controllers to choose wired interrupts over MSIs,
unless implementing a standard like PCI Express that dictates MSIs.

@@

\begin{commentary}
Although an APLIC with multiple interrupt domains may appear to duplicate
the per-source state listed above (source configuration, etc.)\ by a
factor equal to the number of domains, in fact, APLIC implementations
can exploit the fact that each source is ultimately active in only one
domain.
In all domains to which a specific interrupt source has not been
delegated, the state associated with the source appears as read-only
zeros, requiring no physical register bits.
\end{commentary}

@@

If the interrupt domain's harts have IMSICs,
then unless the relevant interrupt
files of those IMSICs support value \z{0x40000000} for register
\z{eidelivery}, setting DM to zero (direct delivery mode) will
have the same effect as setting IE to zero.
See Sections \ref{sec:IMSIC-reg-eidelivery}
and~\ref{sec:AdvPLIC-directMode-intrDelivery}.

@@

An interrupt source is inactive in the interrupt domain if either the
source is delegated to a child domain (D~=~1) or it is not delegated
(D~=~0) and SM is Inactive.
Whenever interrupt source~$i$ is inactive in an interrupt domain,
the corresponding interrupt-pending and interrupt-enable bits within
the domain are read-only zeros, and register \z{target[$i$]} is also
read-only zero.
If source~$i$ is changed from inactive to an active mode, the
interrupt source's pending and enable bits remain zeros, unless set
automatically for a reason specified later in this section or in
Section~\ref{sec:AdvPLIC-pendingBits}, and the defined subfields
of \z{target[$i$]} obtain {\unspecified} values.

@@

It is expected that most systems will arrange the
physical addresses of target IMSICs in a simple
linear correspondence with hart index numbers.
(See Section~\ref{sec:IMSIC-systemMemRegions}.)
Registers \z{mmsiaddrcfg} and \z{mmsiaddrcfgh} (along with
\z{smsiaddrcfg} and \z{smsiaddrcfgh} from the next subsection) allow
sufficiently trusted machine-level software, early after system reset,
to configure the pattern of physical addresses for target IMSICs and
then lock this configuration against subsequent tampering.

@@

If \z{domaincfg}.DM is changed, the \z{target} registers
for all active interrupt sources within the domain obtain
{\unspecified} values in all fields defined for the new delivery mode.

@@

### 2022-09-06 doc/src/MSLevel.tex
Add preface and update status of each chapter
Numerous small improvements to the text

Of the major interrupts controlled by the Privileged Architecture
(numbers 0--15), the Advanced Interrupt Architecture categorizes the
counter overflow interrupt (code~13) as a \emph{local interrupt}.
It is assumed furthermore that any future definitions for
reserved interrupt numbers 14 and 15 will also be local interrupts.
Major interrupt numbers 16--23 and 32--47 are additionally reserved for
standard local interrupts that other {\RISCV} extensions may define.
The remaining major interrupts allocated to the Privileged
Architecture, numbers 0--12, are categorized as not local interrupts.
Taken altogether, Table~\ref{tab:majorIntrCategories}
summarizes the Advanced Interrupt Architecture's categorization
of all major interrupt identities.

>>>>

Of the major interrupts controlled by the Privileged Architecture
(numbers 0--15), this Advanced Interrupt Architecture (AIA) categorizes the
counter overflow interrupt (code~13) as a \emph{local interrupt}.
It is assumed furthermore that any future definitions for
reserved interrupt numbers 14 and 15 will also be local interrupts.
The AIA additionally reserves
major interrupt numbers 16--23 and 32--47 for
standard local interrupts that other {\RISCV} extensions may define.
The remaining major interrupts allocated to the Privileged
Architecture, numbers 0--12, are categorized as not local interrupts.
Taken altogether, Table~\ref{tab:majorIntrCategories}
summarizes the AIA's categorization
of all major interrupt identities.
